---
title: "Mini-Project 2"
author: "Karina Lieb, Alexis Kilayko, and Olivia Baldwin"
date: "March 22, 2018"
output: 
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
#loading the data
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
library(tidyverse)
```

This project looks at data on the 2011-2012 election. We analyze the allocation of contributions from committees. We find out where the most money was being spent, by who, for whom, and what the correlation was between spending and outcome. 

These first 3 graphics show total contributions by state from PACs, organizations, and party organizations committees for the federal election. The states are on the x-axis, and the sum of contributions for a given type of committee is on the y-axis. 

We can see each committee type spent most of their money from different states. Washington, D.C. is the highest contributor to elections for both PACs and PCs. Because this data represents congressional elections, these graphics reveal the influence of DC committees on elections across the country. This is an important aspect of congressional elections: votes made by locals are influenced by people in Washington, who may not share interests. 

Another interesting aspect of these graphs is comparing the states with most contributions between committee types. Most PACs are in DC, so it's intuitive that DC has the most contributions. Virginia is the most-contributing state for organizations, followed by DC, Maryland, and California. California's place is explained by population, but Maryland and Virginia's placements are because of their geographic proximity to DC. While PACs are based in DC, organizations are located primarily in surrounding states. 

The party committee graph also shows a majority of contributions originating from DC, since political parties are based in the capital. The second highest state is Iowa, which didn't show up on top of any of the other graphs. Iowa is the site of the first caucus of the election. Unlike PACs and organizations, parties take interest in winning this initial, precedent-setting caucus. The amount of attention invested in this small state has received criticism for being undemocratic because it amplifies the voices of a small group of people. Our analysis shows that attention isn't the only thing disproportionately given to Iowa: money is, too.  


```{r, warning = FALSE, message = FALSE}
# spending by state (which state's elections are most contributed to?)

#function to see state spending by different types of committees (entity)
entity_picker <- function(entity){

  # wrangling : creating new data frame with relevant information from contributions data frame
  state_spending <- contributions %>%
    filter(entity_type == entity) %>%
    subset(state != '') %>% 
    group_by(state) %>%
    summarise(total_cont = sum(transaction_amt)) %>%
    arrange(desc(total_cont))

  # plotting
  ggplot(data = state_spending, aes(x = reorder(state,-total_cont), y = total_cont)) +
    geom_col() +
    
    theme(axis.text.x = element_text(size = rel(0.9), angle = 80, hjust = 1)) +
  
    ggtitle("Spending By State") +
    labs(x = "State", y = "Total Contribution (million dollars)") +
  
    ggtitle(paste("Total Contributions from", entity, "Committees by State \n(2011-2012 Federal Election Cycle)", sep = " ")) +
    labs(x = "State", y = "Total Contributions ($)") +

    scale_y_continuous(trans = "log2") 
}

entity_picker(entity = "PAC") +
  scale_y_continuous(breaks = c(16, 1024, 65536, 4194304),
                     labels = c("16", "1024", "65536", "4194304"))
entity_picker(entity = "ORG") +
  scale_y_continuous(breaks = c(32, 4096, 524288, 67108864),
                     labels = c("32", "4096", "524288", "67108864"))
entity_picker(entity = "PTY") +
  scale_y_continuous(breaks = c(16, 512, 16384, 524288),
                     labels = c("16", "512", "16384", "524288"))
```

For our second graph, we wanted to examine if there was a correlation between amount of money given to a candidate committee, the candidate's outcome in the election and whether a candidate's incumbency affected the outcome.  To accomplish this, we mapped candidate committee to the x-axis and transaction sum of contributions to the y-axis, using a dot plot.  We mapped outcome to color and we facet wrapped by incumbency.

From this graph, we learn higher contributions don't necessarily correlate to victory and incumbents are more likely to win than non-incumbents.  The results contrasted with our expectation that the more contributed to the committee, the more likely the candidate was to win.  The results did match our expectation that incumbents have a higher frequency of victory due to their experience and visibility among the constituency.

```{r, warning = FALSE, message = FALSE}
# sum of money given to candidate committees

# wrangling : creating new data frame with relevant information from contributions data frame
contributions_org <- contributions %>%
  select(cand_id, transaction_amt) %>%
  group_by(cand_id) %>%
  summarize(transaction_sum = sum(transaction_amt))

# wrangling : creating new data frame with relevant information from house_elections data frame
house_elections_org <- house_elections %>%
  select(fec_id, ge_winner, incumbent) %>%
  rename(cand_id = fec_id)

# wrangling : joining two new data frames
cand_spending <- inner_join(contributions_org, house_elections_org, by = "cand_id")

# wrangling : changing names of observations
cand_spending <- cand_spending %>%
  mutate(new_incumbent = ifelse(incumbent == "TRUE", "Incumbent", "Not Incumbent"),
         new_ge_winner = ifelse(ge_winner == "W", "Win", "Loss"))

# plotting
ggplot(data = cand_spending, aes(x = cand_id, y = transaction_sum, col = new_ge_winner)) +
  geom_point() +
  facet_wrap(~new_incumbent) +
  theme(axis.text.x = element_blank()) +

  ggtitle("Sum of Money Given to House Candidate Committees \n(2011-2012 Federal Election Cycle)") +
  labs(x = "Candidate Committee", y = "Transaction Sum (millions of dollars)", col = "Loss or Win in General Election") +

  scale_y_continuous(breaks = c(0e+00, 2e+06, 4e+06, 6e+06),
                     labels = c("0", "2", "4", "6"))
```

For our final graph, we wanted to examine if there is a trend with categorization of the highest campaign donors during the election and percent of all contributing donors the entity types make up. To view this data, we mapped entity types to x-axis, donation sum to y-axis, and percentage to a gradient.

From this bar graph, we learn that invalid and candidates entities donated the most during the election. Although the most groups who donate are candidates, invalids that represent few of the donation groups donate ten times more money than any others of similar standing. The implications are some liberties are being taken with the FEC's guidelines, for there are limits on how much certain entities can contribute to campaigns. If they don't conform to categories, there is no limit on their involvement. 

```{r, warning = FALSE, message = FALSE}
# name and types/designations of the highest-contributing committees

# write a function for the graphical representation of this!
contributions_by <- function(govt_entity) {
# wrangling relevant variables from committee
committee_info <- committees %>%
  select(cmte_id, cmte_dsgn, cmte_type, org_type, cand_id)

# wrangling relevant data from contributions
contribution_info <- contributions %>%
  select(cmte_id, entity_type, transaction_amt, other_id, cand_id)

# wrangling data from house elections
house_elections_info <- house_elections %>%
  select(fec_id, ge_winner, incumbent, party) %>%
  rename(cand_id = fec_id)

# joining observations of contributions and committees
contri_commit <- committee_info %>%
  full_join(contribution_info, by = c("cmte_id", "cand_id"))

# joining election info
contri_commit <- contri_commit %>%
  right_join(house_elections_info, by = "cand_id")

# data for graph
mutable_data <- contri_commit %>%
  filter(!is.na(entity_type)) %>%
  group_by(entity_type) %>%
  filter(str_detect(cand_id, govt_entity))

mutable_data <- mutable_data %>%
  mutate(entity_count = n(),
         percent = entity_count/nrow(mutable_data)*100,
         transaction_sum = sum(transaction_amt)) %>%
  arrange(desc(transaction_sum))

# graphing it
ggplot(mutable_data, aes(x = reorder(entity_type, -transaction_sum), y = transaction_sum, fill = percent)) +
  geom_bar(stat = "identity", position = position_dodge(), width  = 0.7) +
  theme(axis.text.x = element_text(size = rel(0.9), angle = 7.5, hjust = 1)) +
  scale_y_continuous(trans = "log2")
}

# making a plot by House candidates
plot_3 <- contributions_by(govt_entity = "H")

# bells and whistles
plot_3 +
  ggtitle("Contribution to House Candidates by Entity Type\n(2011-2012 Federal Election Cycle)") +
  labs(x = "Entity Type", y = "Net Amount of Donations by Entities ($)", fill = "Percentage\nof Entities") +
  scale_x_discrete(labels = c("Inv.", "Cand.", "Cand. Cmte.", "Cmte.", "Indv.", "Org.", "Pol. Actn. Cmte.", "Pty. Org.")) +
  scale_y_continuous(labels = c("32", "4096", "524288", "67108864")) +
  scale_fill_gradient(low = "#b6f9dc", high = "#01693c",
                        breaks = c(25,50,75),
                        labels = c("25%", "50%", "75%"))
```



^1^: Link to GitHub repository: https://github.com/akilayko/sds192-mp2
